\begin{theorem}\label{thm:collatz_conjecture}\lean{collatz_conjecture}\leanok
For every natural number $n$, either $n = 0$ or there exists a natural number $k$ such that
$C^k(n) = 1$.
\end{theorem}
\begin{proof}
No proof exists.
\end{proof}

\begin{lemma}\label{lemma:cycle_implies_not_collatz}\lean{cycle_implies_not_collatz}\leanok
If there exists a natural number $n > 4$ and $k \ge 1$ such that $C^k(n) = n$, then the Collatz
conjecture is false {\bf (No-cycle condition)}.
\end{lemma}
\begin{proof}\leanok
Suppose for the sake of contradiction that the Collatz conjecture is true, which would imply every positive natural number eventually reaches $1$.
Since $n > 4 \ge 1$, there exists some $j \ge 0$ such that $C^j(n) = 1$.
Because $n$ is part of a cycle of length $k \ge 1$, we have $C^{j \cdot k}(n) = n$.
Since $j \cdot k \ge j$, we can write $j \cdot k = (j \cdot k - j) + j$, allowing us to decompose the iteration:
$n = C^{j \cdot k}(n) = C^{j \cdot k - j}(C^j(n)) = C^{j \cdot k - j}(1)$.
However, it is a known property of the Collatz map that iterating starting from $1$ only yields values in the set $\{1, 2, 4\}$.
Thus, $C^{j \cdot k - j}(1) \le 4$, which implies $n \le 4$.
This contradicts our initial assumption that $n > 4$, so the conjecture must be false.
\end{proof}

\begin{lemma}\label{lemma:unbounded_orbit_implies_not_collatz}\lean{unbounded_orbit_implies_not_collatz}\leanok
If there exists a natural number $n$ such that its orbit under $C$ is unbounded, then the Collatz
conjecture is false {\bf (No-unbounded-orbit condition)}.
\end{lemma}
\begin{proof}\leanok
Suppose for the sake of contradiction that the Collatz conjecture is true, meaning every positive integer reaches $1$.
For our starting value $n \ge 1$, there must exist some index $j \ge 0$ such that $C^j(n) = 1$.
Let $M$ be the maximum value reached in the first $j$ steps of the orbit, that is, $M = \max_{0 \le i \le j} C^i(n)$.
Because the orbit of $n$ is unbounded, there exists some index $k$ such that $C^k(n) > M + 4$.
If we had $k \le j$, then $C^k(n) \le M < M + 4$, which is a contradiction.
Therefore, we must have $k > j$.
We can then decompose the iteration as $C^k(n) = C^{k-j}(C^j(n)) = C^{k-j}(1)$.
Since iterating the Collatz map from $1$ only produces values in $\{1, 2, 4\}$, we have $C^{k-j}(1) \le 4$.
This implies $C^k(n) \le 4$, which contradicts $C^k(n) > M + 4 \ge 4$.
Thus, the assumption that the conjecture is true must be false.
\end{proof}

\begin{lemma}\label{lemma:bounded_no_cycle_implies_collatz}\lean{bounded_no_cycle_implies_collatz}\leanok
If no natural number $n > 4$ lies on a nontrivial cycle and every orbit under $C$ is bounded, then
the Collatz conjecture is true.
\end{lemma}
\begin{proof}\leanok
Let $m$ be an arbitrary natural number. If $m = 0$, the conjecture trivially holds.
Assume $m \ge 1$. By the boundedness assumption, there exists some bound $B$ such that $C^k(m) \le B$ for all $k \ge 0$.
Because the infinite sequence $(C^k(m))_{k=0}^\infty$ only takes values in the finite set $\{1, 2, \ldots, B\}$, the Pigeonhole Principle implies that there must be a collision.
That is, there exist indices $i < j$ such that $C^i(m) = C^j(m)$. Let $c = C^i(m)$.
Then $C^{j-i}(c) = C^j(m) = C^i(m) = c$, meaning $c$ is part of a cycle of length $j-i \ge 1$.
By our no-cycle assumption, no number strictly greater than $4$ can be part of a cycle, so we must have $c \le 4$.
Since $m \ge 1$, all of its iterates are strictly positive, giving $c \in \{1, 2, 3, 4\}$.
We can verify that every number in this set eventually reaches $1$:
$C^0(1) = 1$, $C^1(2) = 1$, $C^7(3) = 1$ (since $3 \mapsto 10 \mapsto 5 \mapsto 16 \mapsto 8 \mapsto 4 \mapsto 2 \mapsto 1$), and $C^2(4) = 1$.
In all cases, there exists some $j'$ such that $C^{j'}(c) = 1$.
Therefore, starting from $m$, after $i$ steps we reach $c$, and after $j'$ more steps we reach $1$, yielding $C^{i+j'}(m) = 1$.
Thus, the Collatz conjecture is true.
\end{proof}
